Nanophotonic Computational Design

### 2013-03-22 paper/paper.tex
Added a bunch of things, paper ready for submission.

\begin{thebibliography}{99}
% \bibitem{fibergrating} Y. Tang, Z. Wang, L. Wosinski, U. Westergren, and S. He,
%     ``Highly efficient nonuniform grating coupler for silicon-on-insulator 
%     nanophotonic circuits,''
%     Opt. Lett. \textbf{35}, 1290-1292 (2010).
% \bibitem{ridge}  K. K. Lee, D. R. Lim, L.C. Kimerling, J. Shin, and F. Cerrina, 
%     ``Fabrication of ultralow-loss Si/SiO2 waveguides by roughness reduction,''
%     Opt. Lett. \textbf{26}, 1888-1890 (2001).
% \bibitem{pcslow} Y. A. Vlasov, M. O'Boyle, H. F. Hamann, and S. J. McNab,
%     ``Active control of slow light on a chip with photonic crystal waveguides,''
%     Nature \textbf{438}, 65-69 (2005).
% \bibitem{slotfocus} M. Lipson, 
%     ``Guiding, modulating, and emitting light on 
%     silicon-challenges and opportunities,'' 
%     J. Lightwave Technol. \textbf{23}, 4222-4238 (2005).
% \bibitem{active} J. Van Campenhout, P. Rojo Romeo, P. Regreny, C. Seassal, 
%     D. Van Thourhout, S. Verstuyft, L. Di Cioccio, J.-M. Fedeli, 
%     C. Lagahe, and R. Baets, 
%     ``Electrically pumped InP-based microdisk lasers integrated with a 
%     nanophotonic silicon-on-insulator waveguide circuit,'' 
%     Opt. Express \textbf{15}, 6744-6749 (2007).
% \bibitem{metallic} L. Tang, S. E. Kocabas, S. Latif, A. K. Okyay, 
%     D. S. Ly-Gagnon, K. C. Saraswat, and D. A. B. Miller, 
%     ``Nanometre-scale germanium photodetector enhanced by a 
%     near-infrared dipole antenna,'' 
%     Nature Photonics \textbf{2}, 226-229 (2008).
% % \bibitem{fwadia} V. R. Almeida, R. R. Panepucci, and M. Lipson, 
% %     ``Nanotaper for compact mode conversion,'' 
% %     Opt. Lett. \textbf{28}, 1302-1304 (2003) 
% % \bibitem{wwadia} S. G. Johnson, P. Bienstman,  M. A. Skorobogatiy, 
% %     M. Ibanescu1, E. Lidorikis, and J. D. Joannopoulos,
% %     ``Adiabatic theorem and continuous coupled-mode theory for 
% %     efficient taper transitions in photonic crystals,''
% %     Phys. Rev. E \textbf{66}, 066608 (2002)
% % \bibitem{deriv} F. Wang, J. S. Jensen, O. Sigmund, 
% %     ``Robust topology optimization of photonic crystal waveguides with 
% %     tailored dispersion properties.'' 
% %     J. Opt. Soc. Am. B \textbf{28}, 387-397 (2011)
% % \bibitem{boydbook} S. Boyd, and L. Vandenberghe, 
% %     \emph{Convex Optimization} 
% %     (Cambridge University Press, 2004)
% \bibitem{prevwork} J. Lu, S. Boyd, and J. Vuckovic, 
%     ``Inverse design of a three-dimensional nanophotonic resonator,''
%     Opt. Express \textbf{19}, 10563-10750 (2011). 
% \bibitem{veronis} G. Veronis, and S. Fan, 
%     ``Theoretical investigations of compact couplers between dielectric slab 
%     waveguides and two-dimensional metal-dielectric-metal plasmonic waveguides,'' 
%     Opt. Express \textbf{15}, 1211-1221 (2007).
% \bibitem{yang} R. Yang, R. A. Wahsheh, Z. Lu, and M. A. G. Abushagur, 
%     ``Efficient light coupling between dielectric slot waveguide and plasmonic
%     slot waveguide,'' Opt. Lett. \textbf{35}, 649-651 (2010).
% \bibitem{code} \url{www.github.com/JesseLu/objective-first}
\end{thebibliography}

@@

Opening the design space to include many more parameters
    allows for smaller footprint, higher performance devices by definition;
    since original designs are still included in this parameter space.
Unfortunately, the lack of intuition for what such designs might look like and
    the inability to manually search such a large parameter space
    have greatly hindered the ability to employ 
    anything even close to the available parameter space
    for designing nanophotonic components.

@@

Lastly, since all our results are produced by simply specifying
    the functionality and performance of the desired device,
    our results suggest that our method may indeed by able
    to design \emph{all} linear nanophotonic devices.

@@

    In contrast to typical schemes for optimizing physical structures,
        our formulation actually allows for non-zero physics residuals;
        which can be deduced since $A_i(z)x_i-b_i=0$ is not a hard constraint.
    Instead, this formulation is what we call an \emph{objective-first} % Ref
        formulation in that the \emph{design objective} (explained below)
        is prioritized above satisfying physics.

@@

\item 
    The final term in \eq{problem}, $z_\text{min} \le z \le z_\text{max}$,
        is the structure design objective.
    It is used as a relaxation of the binary constraint,
        $z \in {z_\text{min}, z_\text{max}}$,
        which would ensure that the final design be composed 
        of two discrete materials.
\end{enumerate}

\section{Method of solution}
We employed the alternating directions method of multipliers (ADMM) algorithm % Ref!
    in order to solve \eq{problem}.
The ADMM algorithm solves \eq{problem} by iteratively solving for 
    $x_i$, $z$, and a dual variable $u_i$.

Since we are working in three dimensions, solving \eq{problem} for $x_i$ 
    is non-trivial in that it involves millions of variables and
    requires solving for the ill-conditioned $A_i(z)$ matrix.
For this reason, we use a home-built
    finite-difference frequency-domain (FDFD) solver which 
    implements a hardware-accelerated iterative solver % Ref!
    on Amazon's Elastic Compute Cloud.
Critically, our cloud-based solver allows us to scale to solve problems
    with arbitrarily-large number of modes,
    with no significant penalty in runtime.

@@

Lastly, in order to arrive at fully discrete, manufacturable structures,
    we convert $z$ to a boundary parameterization % Ref!
    and tune our structure using a steepest-descent method.

\section{Results}
We demonstrate the effectiveness of our design method 
    by produce designs for a variety of nanophotonic devices.

All of our results are in three dimensions
    and consist of a 250 nm etched silicon slab
    completely surrounded by silica.
The permittivity values of silicon and silica used
    were $\epsilon_\text{Si} = 12.25$ and $\epsilon_\text{SiO$_2$} = 2.25$
    respectively.

Many, if not all, of the produced designs exhibit 
    novel functionality, high efficiency ($>80\%$), and 
    very compact footprints of only a few square vacuum wavelengths;
    while still remaining manufacturable.
We also show that many devices can be designed
    to exhibit different functionality for different input excitations.
Additionally, we show that devices can be designed with large tolerances for
    errors in wavelength, temperature and fabrication.

@@

Our first devices consisted of waveguide mode converters.
Such devices are simple in that they are single-input, single-output devices.
At the same time, such a device is significant because 
    it demonstrates the feasibility of muli-mode on-chip optical networks
    by showing that high-efficiency mode conversion 
    can readily be achieved in planar on-chip nanophotonic structures.

We show, through the design of mode converters for both the TE- and TM-polarized
    waveguide modes, 
    that our method is indeed fully three-dimensional.
Additionally, the devices also demonstrate very small device footprints 
    ($1.6 \times 2.4$ microns for this device in particular).

\subsubsection{TE mode converter}
Our first result was a mode conversion device operating in the TE polarization,
    which we define as that mode where the primary E-field component
    of the waveguide mode is polarized in the plane of the structure.

Our performance specification (\fig{3Db_wgc_te}) 
    for the device was for 90\% to 100\% of the 
    input power to be transferred from the fundamental waveguide mode,
    to the second-order waveguide mode.
At the same time, we specified that no more than 1\% of the input power
    was to remain in the transmitted fundamental mode.

\myfig{3Db_wgc_te}
    {Perfomance specification of the TE mode converter.
    Input mode is the fundamental TE-polarized mode on the left.
    Primary output mode is the second-order mode on the right.
    Output power in the transmitted fundamental mode on the right 
    above 1\% is to be rejected.
    The structure shown is the final three-dimensional design.
    }

The performance of the device is shown in \fig{takeaway}.
The conversion efficiency into the second-order mode is lower
    than desired (86.4\% as opposed to 90\% to 100\%);
    this may be due to evanescent modes ``interfering'' 
    with the output field overlap calculation.

\myfig{takeaway}
    {Structure and E-field at the central plane of the TE mode converter.
    The conversion efficiency into the second-order mode is 86.4\%,
    while the power into the rejection mode (fundamental) is 0.7\%.
    Device footprint is $1.6\times2.4$ microns.}

@@

Since our method is design-by-specification, 
    the design of a TM mode converter requires only
    a small modification to the performance specification of the device;
    namely the polarization of the input and output modes (\fig{wgc_tm}).
Specifically, we still design for 
    $\ge 90\%$ conversion into the second-order mode and 
    a $\le 1\%$ allowance for the fundamental mode to be transmitted.

@@

The performance of the device is shown in \fig{wgc_tm}.
The lower conversion efficiency of 76.9\% in contrast to the TE mode converter
    may be attributed to the lower confinement of the TM waveguide modes.
However, good rejection of only 1\% is still achieved.

@@

\subsection{Splitters}
Next, we demonstrate the design of nanophotonic waveguide mode splitters.
Such devices can be used as multiplexors or demultiplexors and
    are the key component in utilizing a single waveguide to transmit 
    multiple optical signals.

@@

The performance specification for each device is simply to convert
    90\% of the input power in a particular input mode into
    either one of the output modes.
At the same time, we specify that the transmission into the other output mode 
    be kept below 1\% of input power.

\subsubsection{Spatial mode splitter}
We demonstrate what is, to our knowledge, 
    the first design for a three-dimensional
    nanophotonic spatial mode splitter.
Such a device is the key enabler for multi-mode on-chip optical circuits,
    and we show here that they can be designed to be highly efficient
    while utilizing a very small device footprint ($2.8\times2.8$ microns). 
The performance specification is shown in \fig{3Db_spl_modal},
    and the final results is shown in \fig{spl_modal}.

@@

\subsubsection{TE/TM splitter}
In addition to splitting different spatial modes, 
    we show that different polarization can also be split.
\Fig{3Db_spl_tetm} shows the performance specification of a 
    device which is able to separate fundamental TE- and TM-polarized
    waveguide modes into separate arms.
The final, verified result is shown in \fig{spl_tetm}.

@@

\myfig{3Db_spl_tetm}
    {TE/TM splitter performance specification.
    Input mode is either the fundamental TE- or TM-polarized 
        mode on the left.
    Output power into the desired output arm is specified to be greater than 90\%,
        while power into the opposing arm is set to below 1\%.}
\myfig{spl_tetm}
    {TE/TM splitter final result.
    The conversion efficiencies into the upper and lower output arms
        are 87.6\% and 88.8\% respectively, 
        while the rejection powers for the same modes are 1.06\% and 0.58\%.
    Device footprint is $2.8\times2.8$ microns.}

\subsubsection{Wavelength splitter}
Traditional wavelength splitting devices can also be designed using our method.
Here, we show that the 1550 nm and 1310 nm wavelengths can be split
    in a very small device footprint ($2.8\times2.8$ microns).
The performance specification is shown in \fig{3Db_spl_wdm}, 
    and the final result is shown in \fig{spl_wdm}.
\myfig{3Db_spl_wdm}
    {Wavelength splitter performance specification.
    Input mode is the fundamental TE-polarized mode on the left at 
        a wavelength of either 1550 nm or 1330 nm.
    Output modes are the fundamental waveguide modes of either output 
        waveguide on the right.
    Output power into the desired output arm is specified to be greater than 90\%,
        while power into the opposing arm is set to below 1\%.}
\myfig{spl_wdm}
    {Wavelength splitter final result.
    The conversion efficiencies into the upper and lower output arms
        are 83.2\% and 78.7\% respectively, 
        while the rejection powers for the same modes are 0.49\% and 1.66\%.
    Device footprint is $2.8\times2.8$ microns.}

@@

\myfig{3Db_hub_3x3}
    {3$\times$3 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode.
    Output power into the desired output arm is specified to be greater than 90\%,
        no rejection modes are used for computational efficiency.}
\myfig{hub_3x3}
    {3$\times$3 hub final result.
    The conversion efficiencies into the selected output arms
        are 88.6\%, 90.6\%, and 87.3\% for 
        input arms 1, 2, and 3 respectively (top to bottom).}

@@

\myfig{3Db_hub_4x4}
    {4$\times$4 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode.
    Output power into the desired output arm is specified to be greater than 90\%,
        no rejection modes are used for computational efficiency.}
\myfig{hub_4x4_comb}
    {4$\times$4 hub final result.
    The conversion efficiencies into the selected output arms
        are 85.9\%, 88.1\%, 85.4\%, and 84.3\% for
        input arms 1, 2, 3, and 4 respectively (top to bottom).}

@@

Our performance specification (\fig{3Db_hub_2x2x2}) is
    to cross-couple the waveguides at the 1310 nm wavelength,
    but to uncouple the waveguides at the 1550 nm wavelength.
The final result is shown in \fig{hub_2x2x2}.
\myfig{3Db_hub_2x2x2}
    {2$\times$2$\times$2 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode
        at either 1550 nm or 1310 nm wavelengths.
    Output power into the desired output arm is specified to be greater than 80\%,
        no rejection modes are used for computational efficiency.}
\myfig{hub_2x2x2}
    {2$\times$2$\times$2 hub final result.
    The conversion efficiencies at the 1550 nm wavelength 
        are 77.6\% and 73.7\% respectively for the top and bottom inputs.
    At 1310 nm, the respective efficiencies are 75.7\% and 75.2\%.}

\subsection{Fiber couplers}
The capabilities of our method are further demonstrated
    in the design of nanophotonic fiber couplers,
    which couple light from an optical fiber at normal incidence
    into an in-plane waveguide.

@@

\myfig{3Db_fib_pol}
    {Mode-splitting fiber coupler performance specification.
    Input mode consists of the fundamental fiber mode or 
        the third-order, circularly polarized fiber mode.
    Output mode is the fundamental TE-polarized mode of either in-plane waveguide.
    Output power into the desired output arm is specified to be greater than 90\%.}
\myfig{fib_pol}
    {Mode-splitting fiber coupler final result.
    The conversion efficiency for the fundamental fiber mode input is 32.6\% 
        (top plot).
    The conversion efficiency for the third-order fiber mode input is 22.7\% 
        (bottom plot).}

@@

\myfig{3Db_fib_wdm}
    {Wavelength-splitting fiber coupler performance specification.
    Input mode consists of the $E_y$-polarized fundamental fiber mode at either
        the 1310 nm or 1550 nm wavelengths.
    Output mode is the fundamental TE-polarized mode of either in-plane waveguide.
    Output power into the desired output arm is specified to be greater than 90\%.}
\myfig{fib_wdm}
    {Wavelength-splitting fiber coupler final result.
    The conversion efficiency into the in-plane waveguide mode at 1550 nm is 31.6\%.
    The conversion efficiency into the in-plane waveguide mode at 1310 nm is 28.6\%.}

@@

In order to design a broadband wavelength splitter,
    we modify our performance specification to include multiple target 
    wavelengths (with identical desired performance) around the 
    original target wavelengths, as seen in \fig{analysis_top}
    which reveals that broadband operation has been achieved.
The final result for the broadband wavelength splitter is shown in \fig{top_wdm}.
\myfig{analysis_top}
    {Broadband analysis of broadband wavelength splitter.
    The addition of multiple target wavelengths (vertical arrows)
    allows for 
    high-efficiency operation is achieved across a wide bandwidth.}
\myfig{top_wdm}
    {Broadband wavelength splitter final result.
    The efficiencies at the central target wavelengths of 1550 nm and 1310 nm
    exceed those of its narrowband counterpart (\fig{spl_wdm}).}

@@

This result is significant in that it demonstrates
    that the design of broadband devices
    seems to be a valid heuristic in the search for 
    devices which are tolerant to temperature shifts and fabrication error.
Note, however, that our method, as formulated, is also able to
    deal with temperature and fabrication shifts explicitly as well;
    although such results are not demonstrated here.

>>>>

\begin{thebibliography}{99}
\bibitem{ob1_wg} J. Lu, J. Vuckovic, ``Objective-first design of high-efficiency, small-footprint couplers between arbitrary nanophotonic waveguide modes,'' 
    Opt. Express \textbf{20}, 7221-7236 (2012)
\bibitem{admm} S. Boyd, N. Parikh, E. Chu, B. Peleato, and J. Eckstein,
    ``Distributed Optimization and Statistical Learning via the Alternating Direction Method of Multipliers,''
    Foundations and Trends in Machine Learning \textbf{3}, 1-122 (2011). 
\bibitem{wonseok} W. Shin and S. Fan, ``Choice of the perfectly matched layer boundary condition for frequency-domain Maxwell’s equations solvers,''
    J. of Comp. Phys \textbf{231}, 3406-3431 (2012).
\bibitem{levelset} S. Osher and R. Fedkiw, \emph{Level Set Methods and Dynamic Implicit Surfaces: 1st Edition} (Springer, 2002).
\bibitem{2dsplitter} Y. Jiao, S. Fan, and D. A. B. Miller, 
    ``Demonstrations of systematic photonic crystal design and optimization by low rank adjustment: an extremely compact mode separator'', Opt. Lett. \textbf{30}, 140-142 (2005).
\bibitem{baets}F. Van Laere, G. Roelkens, M. Ayre, J. Schrauwen, D. Taillaert, D. Van Thourhout, T. F. Krauss, and R. Baets,
    ``Compact and highly efficient grating couplers between optical fiber and nanophotonic waveguides,''
    J. of Lightwave Tech. \textbf{25}, 151-156 (2007).
% \bibitem{fibergrating} Y. Tang, Z. Wang, L. Wosinski, U. Westergren, and S. He,
%     ``Highly efficient nonuniform grating coupler for silicon-on-insulator 
%     nanophotonic circuits,''
%     Opt. Lett. \textbf{35}, 1290-1292 (2010).
% \bibitem{ridge}  K. K. Lee, D. R. Lim, L.C. Kimerling, J. Shin, and F. Cerrina, 
%     ``Fabrication of ultralow-loss Si/SiO2 waveguides by roughness reduction,''
%     Opt. Lett. \textbf{26}, 1888-1890 (2001).
% \bibitem{pcslow} Y. A. Vlasov, M. O'Boyle, H. F. Hamann, and S. J. McNab,
%     ``Active control of slow light on a chip with photonic crystal waveguides,''
%     Nature \textbf{438}, 65-69 (2005).
% \bibitem{slotfocus} M. Lipson, 
%     ``Guiding, modulating, and emitting light on 
%     silicon-challenges and opportunities,'' 
%     J. Lightwave Technol. \textbf{23}, 4222-4238 (2005).
% \bibitem{active} J. Van Campenhout, P. Rojo Romeo, P. Regreny, C. Seassal, 
%     D. Van Thourhout, S. Verstuyft, L. Di Cioccio, J.-M. Fedeli, 
%     C. Lagahe, and R. Baets, 
%     ``Electrically pumped InP-based microdisk lasers integrated with a 
%     nanophotonic silicon-on-insulator waveguide circuit,'' 
%     Opt. Express \textbf{15}, 6744-6749 (2007).
% \bibitem{metallic} L. Tang, S. E. Kocabas, S. Latif, A. K. Okyay, 
%     D. S. Ly-Gagnon, K. C. Saraswat, and D. A. B. Miller, 
%     ``Nanometre-scale germanium photodetector enhanced by a 
%     near-infrared dipole antenna,'' 
%     Nature Photonics \textbf{2}, 226-229 (2008).
% % \bibitem{fwadia} V. R. Almeida, R. R. Panepucci, and M. Lipson, 
% %     ``Nanotaper for compact mode conversion,'' 
% %     Opt. Lett. \textbf{28}, 1302-1304 (2003) 
% % \bibitem{wwadia} S. G. Johnson, P. Bienstman,  M. A. Skorobogatiy, 
% %     M. Ibanescu1, E. Lidorikis, and J. D. Joannopoulos,
% %     ``Adiabatic theorem and continuous coupled-mode theory for 
% %     efficient taper transitions in photonic crystals,''
% %     Phys. Rev. E \textbf{66}, 066608 (2002)
% % \bibitem{deriv} F. Wang, J. S. Jensen, O. Sigmund, 
% %     ``Robust topology optimization of photonic crystal waveguides with 
% %     tailored dispersion properties.'' 
% %     J. Opt. Soc. Am. B \textbf{28}, 387-397 (2011)
% % \bibitem{boydbook} S. Boyd, and L. Vandenberghe, 
% %     \emph{Convex Optimization} 
% %     (Cambridge University Press, 2004)
% \bibitem{prevwork} J. Lu, S. Boyd, and J. Vuckovic, 
%     ``Inverse design of a three-dimensional nanophotonic resonator,''
%     Opt. Express \textbf{19}, 10563-10750 (2011). 
% \bibitem{veronis} G. Veronis, and S. Fan, 
%     ``Theoretical investigations of compact couplers between dielectric slab 
%     waveguides and two-dimensional metal-dielectric-metal plasmonic waveguides,'' 
%     Opt. Express \textbf{15}, 1211-1221 (2007).
% \bibitem{yang} R. Yang, R. A. Wahsheh, Z. Lu, and M. A. G. Abushagur, 
%     ``Efficient light coupling between dielectric slot waveguide and plasmonic
%     slot waveguide,'' Opt. Lett. \textbf{35}, 649-651 (2010).
% \bibitem{code} \url{www.github.com/JesseLu/objective-first}
\end{thebibliography}

@@

Opening the design space to include many more parameters
    allows for smaller footprint, higher performance devices by definition,
    since original designs are still included in this parameter space.
Unfortunately, the lack of intuition for what such designs might look like and
    the inability to manually search such a large parameter space
    have greatly hindered the ability to achieve this.

@@

Lastly, all our results are produced by simply specifying
    the functionality and performance of the desired device,
    which suggests that our method may indeed by able
    to design \emph{all} linear nanophotonic devices.

@@

    In contrast to typical schemes for optimizing physical structures,
        our formulation actually allows for non-zero physics residuals;
        which can be deduced since $A_i(z)x_i-b_i=0$ is not a hard constraint.
    Instead, this formulation is what we call an \emph{objective-first} \cite{ob1_wg}
        formulation in that the \emph{design objective} (explained below)
        is prioritized above satisfying physics.

@@

\item 
    The final term in \eq{problem}, $z_\text{min} \le z \le z_\text{max}$,
        is the structure design objective.
    It is used as a relaxation of the binary constraint,
        $z \in \{z_\text{min}, z_\text{max}\}$,
        which would ensure that the final design be composed 
        of two discrete materials.
\end{enumerate}

\section{Method of solution}
We employed the alternating directions method of multipliers (ADMM) algorithm 
    \cite{admm}
    in order to solve \eq{problem}.
The ADMM algorithm solves \eq{problem} by iteratively solving for 
    $x_i$, $z$, and a dual variable $u_i$.

Since we are working in three dimensions, solving \eq{problem} for $x_i$ 
    is non-trivial in that it involves millions of variables and
    requires solving for the ill-conditioned $A_i(z)$ matrix.
For this reason, we use a home-built
    finite-difference frequency-domain (FDFD) solver which 
    implements a hardware-accelerated iterative solver\cite{wonseok}
    on Amazon's Elastic Compute Cloud.
Critically, our cloud-based solver allows us to scale to solve problems
    with arbitrarily-large number of modes,
    with no significant penalty in runtime.

@@

Lastly, in order to arrive at fully discrete, manufacturable structures,
    we convert $z$ to a boundary parameterization\cite{levelset}
    and tune our structure using a steepest-descent method.

\section{Results}
We demonstrate the effectiveness of our design method 
    by producing designs for a variety of nanophotonic devices.

All of our results are in three dimensions
    and are planar structures, consisting of a 250 nm etched silicon slab
    completely surrounded by silica.
The permittivity values of silicon and silica used
    were $\epsilon_\text{Si} = 12.25$ and $\epsilon_\text{SiO$_2$} = 2.25$
    respectively.

Many, if not all, of the produced designs exhibit 
    novel functionality, high efficiency, and 
    very compact footprints of only a few square vacuum wavelengths,
    while still remaining manufacturable.
We also show that many devices can be designed
    to exhibit different functionality for different input excitations.
Additionally, we show that devices can be designed with large tolerances for
    errors in wavelength, temperature and fabrication.

@@

Our first devices consisted of waveguide mode converters.
Such devices are simple in that they are single-input, single-output devices.
At the same time, such a device is significant because 
    it demonstrates the feasibility of multi-mode on-chip optical networks
    by showing that high-efficiency mode conversion 
    can readily be achieved in planar on-chip nanophotonic structures.

We show, through the design of mode converters for both the TE- and TM-polarized
    waveguide modes, 
    that our method is indeed fully three-dimensional.
Additionally, the devices also demonstrate very small device footprints 
    ($1.6 \times 2.4$ microns for this device in particular
    operating at 1550 nm wavelength).

\subsubsection{TE mode converter}
Our first result is a mode conversion device operating in the TE polarization,
    where the primary E-field component
    of the waveguide mode is polarized in the plane of the structure.

Our performance specification (\fig{3Db_wgc_te2}) 
    for the device was for $\ge90\%$ of the 
    input power to be transferred from the fundamental waveguide mode,
    to the second-order waveguide mode.
At the same time, we specified that no more than 1\% of the input power
    was to remain in the transmitted fundamental mode.

\myfig{3Db_wgc_te2}
    {Perfomance specification of the TE mode converter.
    Input mode is the fundamental TE-polarized mode on the left.
    Primary output mode is the second-order mode on the right.
    Output power in the transmitted fundamental mode 
    should be no more than 1\%.
    The structure shown is the final three-dimensional design
    (the same holds for all the following figures in the article).
%     The x- and y-axes are in the plane of the structure,
%         with the x-axis is the direction of propagation 
%         for the input and output waveguides.
%     As expected, the z-axis is perpendicular to the plane of the structure.
    }

The performance of the device is shown in \fig{takeaway}.
The conversion efficiency into the second-order mode is lower
    than desired (86.4\%). 
Imperfect conversion may be due to evanescent modes ``interfering'' 
    with the output field overlap calculation.

\myfig{takeaway}
    {Structure and E-field at the central plane of the TE mode converter.
    The conversion efficiency into the second-order mode is 86.4\%,
    while the power into the rejection mode (fundamental) is 0.7\%.
    Device footprint is $1.6\times2.4$ microns.
    Operating wavelength is 1550 nm.}

@@

Since our method is design-by-specification, 
    the design of a TM mode converter requires only
    a small modification to the performance specification of the device;
    namely the polarization of the input and output modes (\fig{3Db_wgc_tm}).
Specifically, we still design for 
    $\ge 90\%$ conversion into the second-order mode and 
    a $\le 1\%$ allowance for the fundamental mode to be transmitted.

@@

The performance of the device is shown in \fig{wgc_tm}.
The lower conversion efficiency of 76.9\% in contrast to the TE mode converter
    may be attributed to the lower confinement of the TM waveguide modes 
    in such thin slabs.
However, good rejection of only 1\% is still achieved.

@@

\subsection{Mode splitters}
Next, we demonstrate the design of nanophotonic waveguide mode splitters.
Such devices can be used as multiplexers or demultiplexers and
    are the key component in utilizing a single waveguide to transmit 
    multiple optical signals.

@@

The performance specification for each device is simply to convert
    more than 90\% of the input power in a particular input mode into
    either one of the output modes.
At the same time, we specify that the transmission into the other output mode 
    be kept below 1\% of input power.

\subsubsection{Spatial mode splitter}
We demonstrate what is, to our knowledge, 
    the first design for a three-dimensional
    nanophotonic spatial mode splitter 
    (previous designs were restricted to two dimensions \cite{2dsplitter}).
Such a device is the key enabler for multi-mode on-chip optical circuits,
    and we show here that they can be designed to be highly efficient
    while utilizing a very small device footprint ($2.8\times2.8$ microns). 
The performance specification is shown in \fig{3Db_spl_modal},
    and the final results is shown in \fig{spl_modal}.

@@

\subsubsection{TE/TM splitter}
In addition to splitting different spatial modes, 
    we show that different polarizations can also be split.
\Fig{3Db_spl_tetm} shows the performance specification of a 
    device which is able to separate fundamental TE-polarized ($E_y$ dominant)
    and TM-polarized ($E_z$ dominant)
    waveguide modes into separate arms.
The final, verified result is shown in \fig{spl_tetm}.

@@

\myfig{3Db_spl_tetm}
    {TE/TM splitter performance specification.
    Input mode is either the fundamental TE-polarized ($E_y$ dominant, top left)
    and TM-polarized ($E_z$ dominant, top right).
    Output power into the desired output arm is specified to be greater than 90\%,
        while power into the opposing arm is set to below 1\%.}
\myfig{spl_tetm}
    {TE/TM splitter final result.
    The conversion efficiencies into the upper and lower output arms
        are 87.6\% and 88.8\% respectively, 
        while the rejection powers for the same modes are 1.06\% and 0.58\%.
    Device footprint is $2.8\times2.8$ microns.}

\subsubsection{Wavelength splitter}
Traditional wavelength splitting devices can also be designed using our method.
Here, we show that the 1550 nm and 1310 nm wavelengths can be split
    in a very small device footprint ($2.8\times2.8$ microns).
The performance specification is shown in \fig{3Db_spl_wdm}, 
    and the final result is shown in \fig{spl_wdm}.
\myfig{3Db_spl_wdm}
    {Wavelength splitter performance specification.
    Input mode is the fundamental TE-polarized mode on the left at 
        a wavelength of either 1550 nm or 1330 nm.
    Output modes are the fundamental waveguide modes of either output 
        waveguide on the right;
        however, the 1550 nm wavelength is directed into the top output, 
        while the 1310 nm wavelength is directed into the bottom output.
    Output power into the desired output arm is specified to be greater than 90\%,
        while power into the opposing arm is set to below 1\%.}
\myfig{spl_wdm}
    {Wavelength splitter final result.
    The conversion efficiencies into the upper and lower output arms
        are 83.2\% and 78.7\% respectively, 
        while the rejection powers for the same modes are 0.49\% and 1.66\%.
    Device footprint is $2.8\times2.8$ microns.}

@@

\myfig{3Db_hub_3x3}
    {3$\times$3 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode.
    Output power into the desired output arm is specified to be greater than 90\%,
        no rejection modes are used for computational efficiency.
    This hub directs input power from input ports 1, 2, and 3 (from top to bottom)
    into output ports 2, 3, and 1 respectively.}
\myfig{hub_3x3}
    {3$\times$3 hub final result.
    The conversion efficiencies into the selected output arms
        are 88.6\%, 90.6\%, and 87.3\% for 
        input arms 1, 2, and 3 respectively (top to bottom).}

@@

\myfig{3Db_hub_4x4}
    {4$\times$4 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode.
    Output power into the desired output arm is specified to be greater than 90\%,
        no rejection modes are used for computational efficiency.
    This hub directs input power from input ports 1, 2, 3, and 4
    into output ports 3, 2, 4, and 1 respectively.}
\myfig{hub_4x4_comb}
    {4$\times$4 hub final result.
    The conversion efficiencies into the selected output arms
        are 85.9\%, 88.1\%, 85.4\%, and 84.3\% for
        input arms 1, 2, 3, and 4 respectively (top to bottom).}

@@

Our performance specification (\fig{3Db_hub_2x2x2}) is
    to cross-couple the waveguides at the 1310 nm wavelength,
    but to uncouple the waveguides at the 1550 nm wavelength.
The final result is shown in \fig{hub_2x2x2}.
\myfig{3Db_hub_2x2x2}
    {2$\times$2$\times$2 hub performance specification.
    Input and output modes all consist of the fundamental TE-polarized mode
        at either 1550 nm or 1310 nm wavelengths.
    Output power into the desired output arm is specified to be greater than 80\%,
        no rejection modes are used for computational efficiency.
    This hub directs input arms 1 and 2 into output arms 1 and 2 at 1550 nm,
        but swaps them at 1310 nm.}
\myfig{hub_2x2x2}
    {2$\times$2$\times$2 hub final result.
    The conversion efficiencies at the 1550 nm wavelength 
        are 77.6\% and 73.7\% respectively for the top and bottom inputs.
    At 1310 nm, the respective efficiencies are 75.7\% and 75.2\%.}

\subsection{Fiber couplers}
The capabilities of our method are further demonstrated
    in the design of nanophotonic fiber couplers,
    which couple light from an optical fiber at normal incidence
    into an in-plane waveguide\cite{baets}.

@@

\myfig{3Db_fib_pol}
    {Mode-splitting fiber coupler performance specification.
    Input mode consists of the fundamental fiber mode or 
        the third-order, circularly polarized fiber mode.
    Output mode is the fundamental TE-polarized mode of either in-plane waveguide.
    The device is designed to couple the fundamental fiber mode into
        the upper output arm, 
        while the third-order fiber mode is coupled into the lower output arm.
    Output power into the desired output arm is specified to be greater than 90\%.}
\myfig{fib_pol}
    {Mode-splitting fiber coupler final result.
    The conversion efficiency for the fundamental fiber mode input is 32.6\% 
        (top plot).
    The conversion efficiency for the third-order fiber mode input is 22.7\% 
        (bottom plot).}

@@

\myfig{3Db_fib_wdm}
    {Wavelength-splitting fiber coupler performance specification.
    Input mode consists of the $E_y$-polarized fundamental fiber mode at either
        the 1310 nm or 1550 nm wavelengths.
    Output mode is the fundamental TE-polarized mode of either in-plane waveguide.
    The structure is designed to guide light at the 1550 nm wavelength
        into the upper output arm,
        while 1310 nm light is guided into the lower output arm.
    Output power into the desired output arm is specified to be greater than 90\%.}
\myfig{fib_wdm}
    {Wavelength-splitting fiber coupler final result.
    The conversion efficiency into the in-plane waveguide mode at 1550 nm is 31.6\%.
    The conversion efficiency into the in-plane waveguide mode at 1310 nm is 28.6\%.}

@@

In order to design a broadband wavelength splitter,
    we modify our performance specification to include multiple target 
    wavelengths (with identical desired performance) around the 
    original target wavelengths, as seen in \fig{analysis_top}
    which reveals that broadband operation has been achieved.
The final result for the broadband wavelength splitter is shown in \fig{top_wdm}.
\myfig{analysis_top}
    {Broadband analysis of broadband wavelength splitter 
        (final design shown in \fig{top_wdm}.
    The addition of multiple target wavelengths (vertical arrows)
    allows for 
    high-efficiency operation is achieved across a wide bandwidth.}
\myfig{top_wdm}
    {Broadband wavelength splitter final result.
    The efficiencies at the central target wavelengths of 1550 nm and 1310 nm
    exceed those of its narrowband counterpart (\fig{spl_wdm}).}

@@

This result is significant in that it demonstrates
    that the design of broadband devices
    seems to be a valid heuristic in the search for 
    devices which are tolerant to temperature shifts and fabrication error.
Note, however, that our method, as formulated, is also able to
    deal with temperature and fabrication shifts explicitly as well,
    although such results are not demonstrated here.

\section{Conclusion}
We have developed and implemented a method to design linear nanophotonic structures
    which are fully three-dimensional and multi-modal,
    have very compact footprints,
    exhibit high efficiency, and
    are manufacturable.
We demonstrate this capability by designing various nanophotonic mode converters,
    splitters, hubs, and fiber couplers.
Critically, many, if not all, of these devices have never been demonstrated before
    and cannot be designed by hand.
In contrast, our method allows user to easily design such devices
    by virtue of our design-by-specification scheme.
